### 2019-06-12 latex/iccv2019/rebuttal/egrebuttal.tex
Start of ICCV rebuttal

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{iccv_rebuttal}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

%%%%%%%%% PAPER ID  - PLEASE UPDATE
\def\iccvPaperID{****} % *** Enter the ICCV Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

\begin{document}

%%%%%%%%% TITLE - PLEASE UPDATE
\title{\LaTeX\ Guidelines for Author Response}  % **** Enter the paper title here

\maketitle
\thispagestyle{empty}

%%%%%%%%% BODY TEXT - ENTER YOUR RESPONSE BELOW
\section{Introduction}
Thank you to all the reviewers for all of your constructive comments and feedback.
We will fix all the minor typos and misunderstandings you have pointed out. 
We hope to point out the more important points here.
% Thank you to all the reviewers for all of your constructive comments and feedback. We fully understand your concerns and we elaborate on them here. We kindly ask you to trust us that the presentation will be clear in the camera-ready and to reward the novelty of the idea in your ratings.

%-------------------------------------------------------------------------
\subsection{Reviewer 1}
\textit{Things are not what they seem}: This is an ominous quote that makes things appear much more dramatic.

%------------------------------------------------------------------------
\subsection{Reviewer 2}
\textit{So it has come to this}: This is another ominous quote that while basically vacuous makes everything appear much more dramatic.

%------------------------------------------------------------------------
\subsection{Reviewer 3}
\textit{So it begins}: This also is an ominous quote that applies to most beginnings but makes things appear much more dramatic.

% \begin{figure}[t]
% \begin{center}
% \fbox{\rule{0pt}{1.8in} \rule{0.9\linewidth}{0pt}}
%    %\includegraphics[width=0.8\linewidth]{egfigure.eps}
% \end{center}
%    \caption{Example of caption.  It is set in Roman so that mathematics
%    (always set in Roman: $B \sin A = A \sin B$) may be included without an
%    ugly clash.}
% \label{fig:long}
% \label{fig:onecol}
% \end{figure}

{\small
\bibliographystyle{ieee}
\bibliography{egbib}
}

\end{document}